---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(knitr)
```

## Project

Does the use of radiation absorbing, RADPAD drapes reduce radiation exposure to cardiac catheterization lab personnel during veterinary interventional procedures?

## Client

Dr. Kursten Pierce and Dr. Kristina Williams, NCSU College of Veterinary Medicine

## Summary

- Interventional cardiology relies on fluoroscopy, medical imaging that utilizes ionizing radiation.
- Radiation exposure poses serious occupational health risks.
- RADPAD drapes have been shown to decrease radiation to lab personnel in human studies.
- The use of RADPAD drapes has not yet been evaluated in veterinary medicine.
- **Study Goal**: if a significant reduction in radiation exposure is shown, RADPAD drapes could be included in standard radiation safety protocol for veterinary procedures.

## First Client Meeting

- Met with both clients last Thursday 5/23
- Obtained detailed explanation of research questions and aims of the study
- Received final data Wednesday 5/29
- Client would like to present summary of analysis at a conference in late July

## Data

```{r}
radpad <- read_csv("RADPAD DATA Stats team 5_28_24_FINAL.xlsx - Sheet1.csv", skip = 1)
```

```{r}
kable(
  radpad |>
    group_by(`RADPAD Drape Used? Y/N`) |>
    summarise(Count = n(),
            "Average Operator Radiation Dose (microSV)" = mean(`Resident 1`))
)
```

- 200 procedures performed on canines
- 6 radiation exposure doses reported per procedure by occupation:
    - Resident 1 (Primary Operator)
    - Resident 2
    - Faculty
    - Tech 1
    - Anesthesia
    - TEE
- Additional variables present:
    - Weight (kg)
    - Age (days)
    - Procedure Type (BPV, PDA PMI, PV Stent)
    - Total Fluoroscopic Time (min)
    
\newpage

## Data Issues

- Radiation exposure dose was not observed for all 6 occupations in every procedure
- Two cases:
    - Personnel not present during the procedure
    - Personnel present but not wearing a radiation badge/badge malfunction

```{r}
nas <- radpad |>
  mutate(across(where(is.character), ~na_if(., "n/a"))) |>
  mutate(across(where(is.character), ~na_if(., "NB")))

nas <-  nas[14:19]

kable(
  nas |>
    summarise(across(everything(), ~sum(is.na(.)))) |>
    pivot_longer(cols = everything(), names_to = "Occupation", values_to = "NA Values") |> 
    arrange(`NA Values`)
)
```

## Analysis Ideas
 
1. ANCOVA Model
    - Treatment factors:
        - RADPAD
        - Occupation
        - Procedure Type
    - Nuisance variables:
        - Weight 
        - Fluoroscopic Time

2. Mixed Model
    - Treatment factors:
        - RADPAD
        - Occupation
    - Random block effect:
        - Procedure
    - Issue with unequal block sizes?
  




