---
title: "The BEAMS Study - Baseline Evaluation And Management of Scatter radiation using RADPAD radiation absorbing drapes"
author: "Analysis completed by Sarah Pagan, Angelice Floyd, and Chandler Ellsworth"
output: pdf_document
header-includes:
  - \usepackage{float}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
#read in relevant packages
library(readxl)
library(tidyverse)
library(lmtest)
library(car)
library(nlme)
library(emmeans)
library(statmod)
library(tweedie)
library(glmmTMB)
library(clubSandwich)
library(sandwich)
#change default print option for tibbles to look nicer when rendering to pdf
print.tbl <- function(x, ...) {
  knitr::kable(x, format = "latex", booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("HOLD_position"))
}
```

```{r}
#read in data
tib <- read_excel("RADPAD_DATA_FINAL_7_3_24.xlsx", skip = 1)
```

```{r}
#select and rename relevant columns
tib <- tib %>% 
  select(`Weight (kg)`, `Procedure  type (BPV, PDA PMI, PV Stent)`, `RADPAD Drape Used? Y/N`, 
         `Faculty`, `Tech 1`, `Resident 1`, `Resident 2`, `TEE`, `Anesthesia`, `Total fluoro time (min)`) %>% 
  rename("Weight" = `Weight (kg)`, "Procedure_Type" = `Procedure  type (BPV, PDA PMI, PV Stent)`, 
         "RADPAD" = `RADPAD Drape Used? Y/N`, "Tech1" = `Tech 1`, "Resident1" = `Resident 1`, 
         "Resident2" = `Resident 2`, , "Time" = `Total fluoro time (min)`) %>%
  mutate(Index = seq(1, nrow(tib), by = 1), Procedure_Type = ifelse(Procedure_Type == "PV stent", "PV Stent", Procedure_Type))
#convert to character type
tib[,c("Index", "Procedure_Type", "RADPAD", "Faculty", "Tech1", "Resident1", "Resident2", "TEE", "Anesthesia")]  <- 
  lapply(tib[,c("Index", "Procedure_Type", "RADPAD", "Faculty", "Tech1", "Resident1", "Resident2", "TEE", "Anesthesia")] , as.character)
#pivot data longer
tib_ <- tib %>%
  pivot_longer(cols = Faculty:Anesthesia, names_to = "Lab_Personnel", values_to = "Dose") %>%
  select(Index, RADPAD, Procedure_Type, Lab_Personnel, Weight, Time, Dose)
#remove relevant data
tib_ <- tib_ %>%
  filter((Dose != "n/a") & (Dose != "NB") & (Dose != "n /a")) %>%
  mutate(Dose = as.numeric(Dose))

#create function to apply the optimal lambda to transform the response variable using Box-Cox
box_transform <- function(response, lambda) {
  if (lambda == 0) {
    response <- log(response)
  } else {
    response <- (response^lambda - 1) / lambda
  }
  return(response)
}

#create function to undo the Box-Cox transformation
back_transform <- function(transformed_response, lambda) {
  if (lambda == 0) {
    # If lambda is zero, use the exponential function
    original_response <- exp(transformed_response)
  } else {
    # Otherwise, use the inverse Box-Cox transformation formula
    original_response <- (transformed_response * lambda + 1)^(1 / lambda)
  }
  return(original_response)
}
```

## Resident 1
```{r, echo = TRUE}
#filter data for only resident 1
res1 <- tib_ %>%
  filter(Lab_Personnel == "Resident1")
#perform Box-Cox transformation on dose
bc <- MASS::boxcox(lm(Dose ~  RADPAD * Procedure_Type + Weight + Time, 
                      data = res1), plotit = FALSE)
#extract the optimal lambda
lambda_optimal <- bc$x[which.max(bc$y)]
#fit model
lm.fit <- lm(box_transform(Dose, lambda_optimal) ~ RADPAD * Procedure_Type + Weight + Time, 
             data = res1)
#analyze anova table
Anova(lm.fit, type = "III")
```
```{r, fig.show = "hold", out.width = "100%", out.height = "46%"}
#look at residual and qq plot
par(mar = c(4, 4, 1.5, 1))
plot(lm.fit, which = c(1,2), sub = "")
```

### Numeric Summaries
```{r}
#calculate summary statistics
summary_stats <- res1 %>% 
  group_by(Procedure_Type, RADPAD) %>%
  summarise(Mean = round(mean(Dose),3), Median = round(median(Dose),3), SD = round(sd(Dose),3))
#print output
print(summary_stats)
#calculate difference and % changes
print(summary_stats %>%
  pivot_wider(names_from = RADPAD, values_from = c(Mean, Median, SD)) %>%
  mutate(
    Mean_Diff = Mean_Y - Mean_N,
    Mean_Percent_Change = round((Mean_Y - Mean_N) / Mean_N * 100, 3),
    Median_Diff = Median_Y - Median_N,
    Median_Percent_Change = round((Median_Y - Median_N) / Median_N * 100, 3)) %>%
  select(Procedure_Type, Mean_Diff, Mean_Percent_Change, Median_Diff, Median_Percent_Change))
```

### Marginal Effect of RADPAD
```{r}
#store Box-Cox transformation for back transformation to response scale
tran <- make.tran("boxcox", bc$x[which.max(bc$y)])
#emmeans on transformed scale
em0 <- emmeans(lm.fit, ~ RADPAD)
c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))

#emmeans on response scale
grid <- ref_grid(lm.fit)
rg <- update(grid, tran = tran)
em <- emmeans(regrid(rg, transform = "response"), ~ RADPAD)
c <- contrast(em, method = "revpairwise", infer = c(TRUE, FALSE))
#combine and print output
print(bind_rows(as_tibble(c0) %>% mutate(Scale = "Box-Cox"), 
          as_tibble(c)  %>% mutate(Scale = "response")))
```

### Simple Effect of RADPAD across Procedure Type
```{r}
#calculate procedure averages
procedure_averages <- res1 %>%
  group_by(Procedure_Type) %>%
  summarise(avg_weight = mean(Weight), avg_time = mean(Time))
#initialize an empty list to store adjusted means
adj_mean_tran <- list()
adj_mean_response <- list()
#initialize an empty list to store contrasts
contrast_tran <- list()
contrast_response <- list()
#loop over each Procedure_Type and calculate the contrasts
for(i in seq_len(nrow(procedure_averages))) {
    proc <- procedure_averages$Procedure_Type[i]
    avg_weight <- procedure_averages$avg_weight[i]
    avg_time <- procedure_averages$avg_time[i]
    
    #create reference grid for current Procedure_Type
    grid <- ref_grid(lm.fit, at = list(Weight = avg_weight, Time = avg_time, Procedure_Type = proc))
    
    #get the estimated marginal means for RADPAD within the current Procedure_Type
    em0 <- emmeans(grid, ~ RADPAD|Procedure_Type)
    # Get the pairwise contrasts for RADPAD
    c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))
    #emmeans on response scale
    rg <- update(grid, tran = tran)
    em <- emmeans(regrid(rg, transform = "response"), ~ RADPAD|Procedure_Type)
    c <- contrast(em, method = "revpairwise", infer = c(TRUE, FALSE))
    #store the adjusted means in the list with the procedure type as the name
    adj_mean_tran[[proc]] <- as_tibble(em0)
    adj_mean_response[[proc]] <- as_tibble(em) %>% rename(emmean = "response")

    #store the contrasts in the list with the procedure type as the name
    contrast_tran[[proc]] <-as_tibble(c0)
    contrast_response[[proc]] <- as_tibble(c)
}
#combine adjusted means and print output
print(bind_rows(adj_mean_tran %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Box-Cox")),
    adj_mean_response %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Response"))) %>%
  mutate_if(is.numeric, round, digits = 4))
#combine contrasts and print output
print(bind_rows(contrast_tran %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Box-Cox")),
    contrast_response %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Response"))) %>%
  mutate_if(is.numeric, round, digits = 4))
```

## Resident 2
```{r, echo = TRUE}
#filter data for only resident 2
res2 <- tib_ %>%
  filter(Lab_Personnel == "Resident2")
#perform Box-Cox transformation on the response variable y
bc <- MASS::boxcox(lm(Dose ~  RADPAD * Procedure_Type + Weight + Time, 
                      data = res2), plotit = FALSE)
#extract the optimal lambda
lambda_optimal <- bc$x[which.max(bc$y)]
#fit model
lm.fit <- lm(box_transform(Dose, lambda_optimal) ~ RADPAD * Procedure_Type + Weight + Time, 
             data = res2)
#analyze anova table
Anova(lm.fit, type = "III")
```

```{r, fig.show = "hold", out.width = "100%", out.height = "46%"}
#look at residual and qq plot
par(mar = c(4, 4, 1.5, 1))
plot(lm.fit, which = c(1,2), sub = "")
```

### Numeric Summaries
```{r}
#calculate summary statistics
summary_stats <- res2 %>% 
  group_by(Procedure_Type, RADPAD) %>%
  summarise(Mean = round(mean(Dose),3), Median = round(median(Dose),3), SD = round(sd(Dose),3))
#print output
print(summary_stats)
#calculate difference and % changes
print(summary_stats %>%
  pivot_wider(names_from = RADPAD, values_from = c(Mean, Median, SD)) %>%
  mutate(
    Mean_Diff = Mean_Y - Mean_N,
    Mean_Percent_Change = round((Mean_Y - Mean_N) / Mean_N * 100, 3),
    Median_Diff = Median_Y - Median_N,
    Median_Percent_Change = round((Median_Y - Median_N) / Median_N * 100, 3)) %>%
  select(Procedure_Type, Mean_Diff, Mean_Percent_Change, Median_Diff, Median_Percent_Change))
```

### Marginal Effect of RADPAD
```{r}
#store Box-Cox transformation for back transformation to response scale
tran <- make.tran("boxcox", bc$x[which.max(bc$y)])
#emmeans on transformed scale
em0 <- emmeans(lm.fit, ~ RADPAD)
c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))

#emmeans on response scale
grid <- ref_grid(lm.fit)
rg <- update(grid, tran = tran)
em <- emmeans(regrid(rg, transform = "response"), ~ RADPAD)
c <- contrast(em, method = "revpairwise", infer = c(TRUE, FALSE))

#combine and print
print(bind_rows(as_tibble(c0) %>% mutate(Scale = "Box-Cox"), 
          as_tibble(c)  %>% mutate(Scale = "response")))
```

### Simple Effect of RADPAD across Procedure Type
```{r}
#calculate procedure averages
procedure_averages <- res2 %>%
  group_by(Procedure_Type) %>%
  summarise(avg_weight = mean(Weight), avg_time = mean(Time))
#initialize an empty list to store adjusted means
adj_mean_tran <- list()
adj_mean_response <- list()
#initialize an empty list to store contrasts
contrast_tran <- list()
contrast_response <- list()
#loop over each Procedure_Type and calculate the contrasts
for(i in seq_len(nrow(procedure_averages))) {
    proc <- procedure_averages$Procedure_Type[i]
    avg_weight <- procedure_averages$avg_weight[i]
    avg_time <- procedure_averages$avg_time[i]
    
    #create reference grid for current Procedure_Type
    grid <- ref_grid(lm.fit, at = list(Weight = avg_weight, Time = avg_time, Procedure_Type = proc))
    
    #get the estimated marginal means for RADPAD within the current Procedure_Type
    em0 <- emmeans(grid, ~ RADPAD|Procedure_Type)
    #get the pairwise contrasts for RADPAD
    c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))
    #emmeans on response scale
    rg <- update(grid, tran = tran)
    em <- emmeans(regrid(rg, transform = "response"), ~ RADPAD|Procedure_Type)
    c <- contrast(em, method = "revpairwise", infer = c(TRUE, FALSE))
    #store the adjusted means in the list with the procedure type as the name
    adj_mean_tran[[proc]] <- as_tibble(em0)
    adj_mean_response[[proc]] <- as_tibble(em) %>% rename(emmean = "response")

    #store the contrasts in the list with the procedure type as the name
    contrast_tran[[proc]] <-as_tibble(c0)
    contrast_response[[proc]] <- as_tibble(c)
}
#combine adjusted means and print
print(bind_rows(adj_mean_tran %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Box-Cox")),
    adj_mean_response %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Response"))) %>%
  mutate_if(is.numeric, round, digits = 4))
#combine contrasts and print
print(bind_rows(contrast_tran %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Box-Cox")),
    contrast_response %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Response"))) %>%
  mutate_if(is.numeric, round, digits = 4))
```

## Faculty
```{r, echo = TRUE}
#filter data for only faculty
Faculty <- tib_ %>%
  filter(Lab_Personnel == "Faculty")
# Perform Box-Cox transformation on the response variable y
bc <- MASS::boxcox(lm(Dose ~  RADPAD * Procedure_Type + Weight + Time, 
                      data = Faculty), plotit = FALSE)
#store Box-Cox transformation for back transformation to response scale
tran <- make.tran("boxcox", bc$x[which.max(bc$y)])
#ordinary least squares
gls.fit <- gls(tran$linkfun(Dose) ~ RADPAD * Procedure_Type + Weight + Time, 
    data = Faculty)
#unequal variance between RADPAD levels
gls.fit2 <- gls(tran$linkfun(Dose) ~ RADPAD * Procedure_Type + Weight + Time, 
    data = Faculty,
    weights = varIdent(form = ~ 1 | RADPAD))
#unequal variance between procedure type levels
gls.fit3 <- gls(tran$linkfun(Dose) ~ RADPAD * Procedure_Type + Weight + Time, 
    data = Faculty,
    weights = varIdent(form = ~ 1 | Procedure_Type))
#unequal variance between combinations of RADPAD and procedure type levels
gls.fit4 <- gls(tran$linkfun(Dose) ~ RADPAD * Procedure_Type + Weight + Time, 
    data = Faculty,
    weights = varIdent(form = ~ 1 | RADPAD*Procedure_Type))
#output AIC for each of the models
df <- AIC(gls.fit, gls.fit2, gls.fit3, gls.fit4)
#change to tibble
tib <- as_tibble(df) %>% 
  mutate(Model = c("Equal Variance", "RADPAD Unequal Variance", "PType Unequal Variance", "RADPAD*PType Unequal Variance")) %>%
  select(Model, df, AIC)
#print tibble
print(tib)
#compare ols and unequal error variance for RADPAD levels with anova
df <- anova(gls.fit, gls.fit2)
#show results
tib <- as_tibble(df)
#change names of formula
tib$call <- c("Equal Variance", "RADPAD Unequal Variance")
#print tibble
print(tib)
#output ANOVA table
Anova(gls.fit2, Type = "III")
```

```{r, fig.show = "hold", out.width = "100%", out.height = "46%"}
#get the residuals (standardized)
residual <- residuals(gls.fit2, type = 'pearson')
#get the fitted values
fitted_values <- fitted(gls.fit2)
#plot residuals vs fitted
plot(fitted_values, residual, xlab = "Fitted values", ylab = "Pearson residuals", main = "Residuals vs Fitted", pch = 1, col = "dodgerblue")
abline(h=0)
#normal QQ plot
qqnorm(residual, pch = 1, col = "dodgerblue")
abline(0,1, lwd=2)
```

### Numeric Summaries
```{r}
#calculate summary statistics
summary_stats <- Faculty %>% 
  group_by(Procedure_Type, RADPAD) %>%
  summarise(Mean = round(mean(Dose),3), Median = round(median(Dose),3), SD = round(sd(Dose),3))
#print output
print(summary_stats)
#calculate difference and % changes
print(summary_stats %>%
  pivot_wider(names_from = RADPAD, values_from = c(Mean, Median, SD)) %>%
  mutate(
    Mean_Diff = Mean_Y - Mean_N,
    Mean_Percent_Change = round((Mean_Y - Mean_N) / Mean_N * 100, 3),
    Median_Diff = Median_Y - Median_N,
    Median_Percent_Change = round((Median_Y - Median_N) / Median_N * 100, 3)) %>%
  select(Procedure_Type, Mean_Diff, Mean_Percent_Change, Median_Diff, Median_Percent_Change))
```

### Marginal Effect of RADPAD
```{r}
#emmeans on transformed scale
em0 <- emmeans(gls.fit2, ~ RADPAD)
c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))

#emmeans on response scale
grid <- ref_grid(gls.fit2)
rg <- update(grid, tran = tran)
em <- emmeans(regrid(rg, transform = "response"), ~ RADPAD)
c <- contrast(em, method = "revpairwise", infer = c(TRUE, FALSE))

#combine and print
print(bind_rows(as_tibble(c0) %>% mutate(Scale = "Box-Cox"), 
          as_tibble(c)  %>% mutate(Scale = "response")))
```

### Simple Effect of RADPAD across Procedure Type
```{r}
#calculate procedure averages
procedure_averages <- Faculty %>%
  group_by(Procedure_Type) %>%
  summarise(avg_weight = mean(Weight), avg_time = mean(Time))
#initialize an empty list to store adjusted means
adj_mean_tran <- list()
adj_mean_response <- list()
#initialize an empty list to store contrasts
contrast_tran <- list()
contrast_response <- list()
#loop over each Procedure_Type and calculate the contrasts
for(i in seq_len(nrow(procedure_averages))) {
    proc <- procedure_averages$Procedure_Type[i]
    avg_weight <- procedure_averages$avg_weight[i]
    avg_time <- procedure_averages$avg_time[i]
    
    #create reference grid for current Procedure_Type
    grid <- ref_grid(gls.fit2, at = list(Weight = avg_weight, Time = avg_time, Procedure_Type = proc))
    
    #get the estimated marginal means for RADPAD within the current Procedure_Type
    em0 <- emmeans(grid, ~ RADPAD|Procedure_Type)
    #get the pairwise contrasts for RADPAD
    c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))
    #emmeans on response scale
    rg <- update(grid, tran = tran)
    em <- emmeans(regrid(rg, transform = "response"), ~ RADPAD|Procedure_Type)
    c <- contrast(em, method = "revpairwise", infer = c(TRUE, FALSE))
    # Store the adjusted means in the list with the procedure type as the name
    adj_mean_tran[[proc]] <- as_tibble(em0)
    adj_mean_response[[proc]] <- as_tibble(em) %>% rename(emmean = "response")

    #store the contrasts in the list with the procedure type as the name
    contrast_tran[[proc]] <-as_tibble(c0)
    contrast_response[[proc]] <- as_tibble(c)
}
#combine adjusted means and print
print(bind_rows(adj_mean_tran %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Box-Cox")),
    adj_mean_response %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Response"))) %>%
  mutate_if(is.numeric, round, digits = 4))
#combine contrasts and print
print(bind_rows(contrast_tran %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Box-Cox")),
    contrast_response %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Response"))) %>%
  mutate_if(is.numeric, round, digits = 4))
```

## TEE
```{r, echo = TRUE}
#filter data for only TEE
TEE <- tib_ %>%
  filter(Lab_Personnel == "TEE")
# Perform Box-Cox transformation on the response variable y
bc <- MASS::boxcox(lm(Dose ~  RADPAD * Procedure_Type + Weight + Time, 
                      data = TEE), plotit = FALSE)
#store Box-Cox transformation for back transformation to response scale
tran <- make.tran("boxcox", bc$x[which.max(bc$y)])
#ordinary least squares
gls.fit <- gls(tran$linkfun(Dose) ~ RADPAD * Procedure_Type + Weight + Time, 
    data = TEE)
#unequal variance between RADPAD levels
gls.fit2 <- gls(tran$linkfun(Dose) ~ RADPAD * Procedure_Type + Weight + Time, 
    data = TEE,
    weights = varIdent(form = ~ 1 | RADPAD))
#unequal variance between procedure type levels
gls.fit3 <- gls(tran$linkfun(Dose) ~ RADPAD * Procedure_Type + Weight + Time, 
    data = TEE,
    weights = varIdent(form = ~ 1 | Procedure_Type))
#unequal variance between combinations of RADPAD and procedure type levels
gls.fit4 <- gls(tran$linkfun(Dose) ~ RADPAD * Procedure_Type + Weight + Time, 
    data = TEE,
    weights = varIdent(form = ~ 1 | RADPAD*Procedure_Type))
#output AIC for each of the models
df <- AIC(gls.fit, gls.fit2, gls.fit3, gls.fit4)
#change to tibble
tib <- as_tibble(df) %>% 
  mutate(Model = c("Equal Variance", "RADPAD Unequal Variance", "PType Unequal Variance", "RADPAD*PType Unequal Variance")) %>%
  select(Model, df, AIC)
#print tibble
print(tib)
#compare ols and unequal error variance for procedure type levels with anova
df <- anova(gls.fit, gls.fit3)
#show results
tib <- as_tibble(df)
#change names of formula
tib$call <- c("Equal Variance", "RADPAD Unequal Variance")
#print tibble
print(tib)
#output ANOVA table
Anova(gls.fit2, Type = "III")
```

```{r, fig.show = "hold", out.width = "100%", out.height = "46%"}
#get the residuals (standardized)
residual <- residuals(gls.fit3, type = 'pearson')
#get the fitted values
fitted_values <- fitted(gls.fit3)
#plot residuals vs fitted
plot(fitted_values, residual, xlab = "Fitted values", ylab = "Pearson residuals", main = "Residuals vs Fitted", pch = 1, col = "dodgerblue")
abline(h=0)
#normal QQ plot
qqnorm(residual, pch = 1, col = "dodgerblue")
abline(0,1, lwd=2)
```

### Numeric Summaries
```{r}
#calculate summary statistics
summary_stats <- TEE %>% 
  group_by(Procedure_Type, RADPAD) %>%
  summarise(Mean = round(mean(Dose),3), Median = round(median(Dose),3), SD = round(sd(Dose),3))
#print output
print(summary_stats)
#calculate difference and % changes
print(summary_stats %>%
  pivot_wider(names_from = RADPAD, values_from = c(Mean, Median, SD)) %>%
  mutate(
    Mean_Diff = Mean_Y - Mean_N,
    Mean_Percent_Change = round((Mean_Y - Mean_N) / Mean_N * 100, 3),
    Median_Diff = Median_Y - Median_N,
    Median_Percent_Change = round((Median_Y - Median_N) / Median_N * 100, 3)) %>%
  select(Procedure_Type, Mean_Diff, Mean_Percent_Change, Median_Diff, Median_Percent_Change))
```

### Marginal Effect of RADPAD
```{r}
#emmeans on transformed scale
em0 <- emmeans(gls.fit3, ~ RADPAD)
c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))

#emmeans on response scale
grid <- ref_grid(gls.fit3)
rg <- update(grid, tran = tran)
em <- emmeans(regrid(rg, transform = "response"), ~ RADPAD)
c <- contrast(em, method = "revpairwise", infer = c(TRUE, FALSE))

#combine and print
print(bind_rows(as_tibble(c0) %>% mutate(Scale = "Box-Cox"), 
          as_tibble(c)  %>% mutate(Scale = "response")))
```

### Simple Effect of RADPAD across Procedure Type
```{r}
#calculate procedure averages
procedure_averages <- TEE %>%
  group_by(Procedure_Type) %>%
  summarise(avg_weight = mean(Weight), avg_time = mean(Time))
#initialize an empty list to store adjusted means
adj_mean_tran <- list()
adj_mean_response <- list()
#initialize an empty list to store contrasts
contrast_tran <- list()
contrast_response <- list()
# Loop over each Procedure_Type and calculate the contrasts
for(i in seq_len(nrow(procedure_averages))) {
    proc <- procedure_averages$Procedure_Type[i]
    avg_weight <- procedure_averages$avg_weight[i]
    avg_time <- procedure_averages$avg_time[i]
    
    # Create reference grid for current Procedure_Type
    grid <- ref_grid(gls.fit3, at = list(Weight = avg_weight, Time = avg_time, Procedure_Type = proc))
    
    # Get the estimated marginal means for RADPAD within the current Procedure_Type
    em0 <- emmeans(grid, ~ RADPAD|Procedure_Type)
    # Get the pairwise contrasts for RADPAD
    c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))
    #emmeans on response scale
    rg <- update(grid, tran = tran)
    em <- emmeans(regrid(rg, transform = "response"), ~ RADPAD|Procedure_Type)
    c <- contrast(em, method = "revpairwise", infer = c(TRUE, FALSE))
    # Store the adjusted means in the list with the procedure type as the name
    adj_mean_tran[[proc]] <- as_tibble(em0)
    adj_mean_response[[proc]] <- as_tibble(em) %>% rename(emmean = "response")

    # Store the contrasts in the list with the procedure type as the name
    contrast_tran[[proc]] <-as_tibble(c0)
    contrast_response[[proc]] <- as_tibble(c)
}
#combine adjusted means and print
print(bind_rows(adj_mean_tran %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Box-Cox")),
    adj_mean_response %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Response"))) %>%
  mutate_if(is.numeric, round, digits = 4))
#combine contrasts and print
print(bind_rows(contrast_tran %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Box-Cox")),
    contrast_response %>%
    imap_dfr(~ .x %>% mutate(Procedure_Type = .y, Scale = "Response"))) %>%
  mutate_if(is.numeric, round, digits = 4))
```

## Anesthesia
```{r, echo = TRUE}
#filter data for only anesthesia
anesthesia <- tib_ %>%
  filter(Lab_Personnel == "Anesthesia")
#fit a GLM with tweedie family
glm.fit <- glm(Dose ~ RADPAD*Procedure_Type + Weight + Time, 
                     family = Gamma(link = "log"), 
                     data = anesthesia)
#look at estimates
Anova(glm.fit, type = "III", white.adjust = TRUE)
```

```{r, fig.show = "hold", out.width = "100%", out.height = "46%"}
#look at diagnostic plots
plot(DHARMa:: simulateResiduals(fittedModel = glm.fit))
#plot deviance residuals versus fitted values
plot(residuals(glm.fit) ~ predict(glm.fit, type = "link"),
     xlab = expression(hat(eta)), ylab = "Deviance residuals")
#qqnorm of deviance residuals
qqnorm(residuals(glm.fit))
qqline(residuals(glm.fit))
```

### Numeric Summaries
```{r}
#calculate summary statistics
summary_stats <- anesthesia %>% 
  group_by(Procedure_Type, RADPAD) %>%
  summarise(Mean = round(mean(Dose),3), Median = round(median(Dose),3), SD = round(sd(Dose),3))
#print output
print(summary_stats)
#calculate difference and % changes
print(summary_stats %>%
  pivot_wider(names_from = RADPAD, values_from = c(Mean, Median, SD)) %>%
  mutate(
    Mean_Diff = Mean_Y - Mean_N,
    Mean_Percent_Change = round((Mean_Y - Mean_N) / Mean_N * 100, 3),
    Median_Diff = Median_Y - Median_N,
    Median_Percent_Change = round((Median_Y - Median_N) / Median_N * 100, 3)) %>%
  select(Procedure_Type, Mean_Diff, Mean_Percent_Change, Median_Diff, Median_Percent_Change))
```

### Marginal Effect of RADPAD
```{r}
em0 <- emmeans(glm.fit, ~ RADPAD|Procedure_Type, vcov. = vcovHC(glm.fit, type = "HC3"))
c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))
c0

#bind_rows(as.data.frame(c0) %>% mutate(Scale = "Box-Cox"), 
          #as.data.frame(c)  %>% mutate(Scale = "resonse"))
```

## GLM on all but Tech1
```{r, echo = TRUE}
#filter data to not include Tech1
noTech <- tib_ %>% filter(Lab_Personnel != "Tech1")
#fit model
glm.fit <- glm(Dose ~ RADPAD*Procedure_Type*Lab_Personnel + Weight + Time, 
                     family = Gamma(link = "log"), 
                     data = noTech)
```

```{r, , fig.show = "hold", out.width = "100%", out.height = "46%"}
#look at diagnostic plots
plot(DHARMa:: simulateResiduals(fittedModel = glm.fit))
#plot deviance residuals versus fitted values
plot(residuals(glm.fit) ~ predict(glm.fit, type = "link"),
     xlab = expression(hat(eta)), ylab = "Deviance residuals")
#qqnorm of deviance residuals
qqnorm(residuals(glm.fit))
qqline(residuals(glm.fit))
```






















