---
title: "RADPAD Data Analysis"
author: "Chandler Ellsworth, Angelice Floyd, and Sarah Pagan"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      out.width = "50%", out.height = "50%")
```

```{r}
## libraries
library(readxl)
library(tidyverse)
library(knitr)
library(kableExtra)
library(car)
library(emmeans)
library(nlme)
library(DHARMa)
```

```{r}
## data processing
radpad <- read_excel("RADPAD_DATA_FINAL_7_3_24.xlsx", skip = 1)
r <- radpad |>
  select(`Weight (kg)`, `Procedure  type (BPV, PDA PMI, PV Stent)`,
         `RADPAD Drape Used? Y/N`, `Faculty`, `Tech 1`, `Resident 1`,
         `Resident 2`, `TEE`, `Anesthesia`, `Total fluoro time (min)`) |>
  mutate(ID = 1:nrow(radpad)) |>
  rename(Weight = `Weight (kg)`) |>
  rename(Procedure_Type = `Procedure  type (BPV, PDA PMI, PV Stent)`) |>
  mutate(Procedure_Type = replace(Procedure_Type,
                                  Procedure_Type == "PV stent",
                                  "PV Stent")) |>
  rename(Time = `Total fluoro time (min)`) |>
  rename(RADPAD = `RADPAD Drape Used? Y/N`) |>
  mutate(`Resident 1` = as.character(`Resident 1`)) |>
  pivot_longer(Faculty:Anesthesia,
               names_to = "Lab_Personnel",
               values_to = "Dose") |>
  filter((Dose != "n/a") & (Dose != "NB")) |>
  mutate(Dose = as.numeric(Dose))
```

```{r}
## subset data by lab personnel
r1 <- r |>
  filter(Lab_Personnel == "Resident 1")
r2 <- r |>
  filter(Lab_Personnel == "Resident 2")
rF <- r |>
  filter(Lab_Personnel == "Faculty")
rTee <- r |>
  filter(Lab_Personnel == "TEE")
rTech <- r |>
  filter(Lab_Personnel == "Tech 1")
rA <- r |>
  filter(Lab_Personnel == "Anesthesia")
```

## Resident 1

### Observed Difference in Median Dose

```{r}
r1 |>
  group_by(Procedure_Type) |>
  summarise(NODPAD = median(ifelse(RADPAD == "N", Dose, NA), na.rm = TRUE),
            RADPAD = median(ifelse(RADPAD == "Y", Dose, NA), na.rm = TRUE)) |>
  mutate(Difference = RADPAD - NODPAD) |> 
  mutate(Perecent = Difference/NODPAD*100) |>
  kable(digits = 2, caption = "Median Dose (microSV) Statistics for Resident 1",
        col.names = c("Procedure", "Median NOPAD", "Median RADPAD",
                      "Difference", "Percent Change"))|>
  kable_styling()
```

### ANCOVA Model Analysis of Variance

```{r}
## fit OLS model
fit_r1 <- lm(log(Dose) ~ RADPAD * Procedure_Type + Weight + Time, data = r1)
Anova(fit_r1, type = "III") |>
  kable(digits = 3, caption = "Resident 1 Analysis of Variance")
```

### Contrasts 

```{r}
## RADPAD by Procedure Type effect on log scale
em0 <- emmeans(fit_r1, ~RADPAD|Procedure_Type)
c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))
data.frame(c0) |>
  kable(digits = 3,
        caption = "Resident 1 RADPAD by Procedure Type Contrast on Log Scale") |>
  kable_styling()

## RADPAD by Procedure Type effect ratio
em <- emmeans(fit_r1, type = "response", ~RADPAD|Procedure_Type)
c1 <- contrast(em, method = "revpairwise", infer = c(TRUE, TRUE))
data.frame(c1) |>
  select(-df) |>
  kable(digits = 3,
        caption = "Resident 1 RADPAD by Procedure Type Contrast, Ratio of Geometric Means") |>
  kable_styling()

## RADPAD main effect on log scale
em2 <- emmeans(fit_r1, ~RADPAD)
c2 <- contrast(em2, "revpairwise", infer = c(TRUE, TRUE))
data.frame(c2) |>
  kable(digits = 3, caption = "Resident 1 RADPAD Contrast on Log Scale") |>
  kable_styling()

## RADPAD main effect ratio
em3 <- emmeans(fit_r1, type = "response", ~RADPAD)
c3 <- contrast(em3, method = "revpairwise", infer = c(TRUE, TRUE))
data.frame(c3) |>
  select(-df) |>
  kable(digits = 3,
        caption = "Resident 1 RADPAD Contrast, Ratio of Geometric Means") |>
  kable_styling()
```

\newpage

## Resident 2

### Observed Difference in Median Dose

```{r}
r2 |>
  group_by(Procedure_Type) |>
  summarise(NODPAD = median(ifelse(RADPAD == "N", Dose, NA), na.rm = TRUE),
            RADPAD = median(ifelse(RADPAD == "Y", Dose, NA), na.rm = TRUE)) |>
  mutate(Difference = RADPAD - NODPAD) |> 
  mutate(Perecent = Difference/NODPAD*100) |>
  kable(digits = 2, caption = "Median Dose (microSV) Statistics for Resident 2",
        col.names = c("Procedure", "Median NOPAD", "Median RADPAD",
                      "Difference", "Percent Change"))|>
  kable_styling()
```

### ANCOVA Model Analysis of Variance

```{r}
## fit OLS model
fit <- lm(Dose ~ RADPAD * Procedure_Type + Weight + Time, data = r2)

## get box cox transformation for Dose
b <- MASS::boxcox(fit, plotit = FALSE)
## make transformation
tran <- make.tran("boxcox", b$x[which.max(b$y)])
## fit model with box cox transformation
fit_r2 <- lm(tran$linkfun(Dose) ~ RADPAD*Procedure_Type + Weight + Time,
             data = r2)
Anova(fit_r2, type = "III") |>
  kable(digits = 3, caption = "Resident 2 Analysis of Variance")
```

### Contrasts

```{r}
## RADPAD by Procedure Type effect on transformed scale
em0 <- emmeans(fit_r2, ~RADPAD|Procedure_Type)
c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))
data.frame(c0) |>
  kable(digits = 3,
        caption = "Resident 2 RADPAD by Procedure Type Contrast on Box-Cox Scale") |>
  kable_styling()

## RADPAD main effect on transformed scale
em2 <- emmeans(fit_r2, ~RADPAD)
c2 <- contrast(em2, "revpairwise", infer = c(TRUE, TRUE))
data.frame(c2) |>
  kable(digits = 3, caption = "Resident 2 RADPAD Contrast on Box-Cox Scale") |>
  kable_styling()
```

\newpage

## Faculty

### Observed Difference in Median Dose

```{r}
rF |>
  group_by(Procedure_Type) |>
  summarise(NODPAD = median(ifelse(RADPAD == "N", Dose, NA), na.rm = TRUE),
            RADPAD = median(ifelse(RADPAD == "Y", Dose, NA), na.rm = TRUE)) |>
  mutate(Difference = RADPAD - NODPAD) |> 
  mutate(Perecent = Difference/NODPAD*100) |>
  kable(digits = 2, caption = "Median Dose (microSV) Statistics for Faculty",
        col.names = c("Procedure", "Median NOPAD", "Median RADPAD",
                      "Difference", "Percent Change"))|>
  kable_styling()
```

### GLS Model Analysis of Variance

```{r}
## fit OLS model
fit <- lm(Dose ~ RADPAD * Procedure_Type + Weight + Time, data = rF)
## get box cox transformation for Dose
b <- MASS::boxcox(fit, plotit = FALSE)
## make transformation
tran <- make.tran("boxcox", b$x[which.max(b$y)])

## fit GLS models under different variance structures with box cox transformation
gls_fit0 <- gls(tran$linkfun(Dose) ~ RADPAD*Procedure_Type + Weight + Time,
                data = rF)
gls_fit1 <- gls(tran$linkfun(Dose) ~ RADPAD*Procedure_Type + Weight + Time,
                data = rF,
                weights = varIdent(form = ~ 1 | RADPAD))
gls_fit2 <- gls(tran$linkfun(Dose) ~ RADPAD*Procedure_Type  + Weight + Time,
                data = rF,
                weights = varIdent(form = ~ 1 | Procedure_Type))
gls_fit3 <- gls(tran$linkfun(Dose) ~ RADPAD*Procedure_Type + Weight + Time,
                data = rF,
                weights = varIdent(form = ~ 1 | RADPAD*Procedure_Type))
## identify best model using AIC criterion
aic_results <- AIC(gls_fit0, gls_fit1, gls_fit2, gls_fit3)
## best model is gls_fit1 with RADPAD unequal variance
fit_rF <- gls_fit1

Anova(fit_rF, type = "III") |>
  kable(digits = 3, caption = "Faculty Analysis of Variance")
```

### Contrasts 

```{r}
## RADPAD by Procedure Type effect on transformed scale
em0 <- emmeans(fit_rF, ~RADPAD|Procedure_Type)
c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))
data.frame(c0) |>
  kable(digits = 3,
        caption = "Faculty RADPAD by Procedure Type Contrast on Box-Cox Scale") |>
  kable_styling()

## RADPAD main effect on transformed scale
em2 <- emmeans(fit_rF, ~RADPAD)
c2 <- contrast(em2, "revpairwise", infer = c(TRUE, TRUE))
data.frame(c2) |>
  kable(digits = 3, caption = "Faculty RADPAD Contrast on Box-Cox Scale") |>
  kable_styling()
```

\newpage

## TEE

### Observed Difference in Median Dose

```{r}
rTee |>
  group_by(Procedure_Type) |>
  summarise(NODPAD = median(ifelse(RADPAD == "N", Dose, NA), na.rm = TRUE),
            RADPAD = median(ifelse(RADPAD == "Y", Dose, NA), na.rm = TRUE)) |>
  mutate(Difference = RADPAD - NODPAD) |> 
  mutate(Perecent = Difference/NODPAD*100) |>
  kable(digits = 2, caption = "Median Dose (microSV) Statistics for TEE",
        col.names = c("Procedure", "Median NOPAD", "Median RADPAD",
                      "Difference", "Percent Change"))|>
  kable_styling()
```

### GLS Model Analysis of Variance

```{r}
## fit OLS model
fit <- lm(Dose ~ RADPAD * Procedure_Type + Weight + Time, data = rTee)
## get box cox transformation for Dose
b <- MASS::boxcox(fit, plotit = FALSE)
## make transformation
tran <- make.tran("boxcox", b$x[which.max(b$y)])

## fit GLS models under different variance structures with box cox transformation
gls_fit0 <- gls(tran$linkfun(Dose) ~ RADPAD*Procedure_Type + Weight + Time,
                data = rTee)
gls_fit1 <- gls(tran$linkfun(Dose) ~ RADPAD*Procedure_Type + Weight + Time,
                data = rTee,
                weights = varIdent(form = ~ 1 | RADPAD))
gls_fit2 <- gls(tran$linkfun(Dose) ~ RADPAD*Procedure_Type  + Weight + Time,
                data = rTee,
                weights = varIdent(form = ~ 1 | Procedure_Type))
gls_fit3 <- gls(tran$linkfun(Dose) ~ RADPAD*Procedure_Type + Weight + Time,
                data = rTee,
                weights = varIdent(form = ~ 1 | RADPAD*Procedure_Type))
## identify best model using AIC criterion
aic_results <- AIC(gls_fit0, gls_fit1, gls_fit2, gls_fit3)
## best model is gls_fit1 with RADPAD unequal variance
fit_rTee <- gls_fit1

Anova(fit_rF, type = "III") |>
  kable(digits = 3, caption = "TEE Analysis of Variance")
```

### Contrasts 

```{r}
## RADPAD by Procedure Type effect on transformed scale
em0 <- emmeans(fit_rTee, ~RADPAD|Procedure_Type)
c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))
data.frame(c0) |>
  kable(digits = 3,
        caption = "TEE RADPAD by Procedure Type Contrast on Box-Cox Scale") |>
  kable_styling()

## RADPAD main effect on transformed scale
em2 <- emmeans(fit_rTee, ~RADPAD)
c2 <- contrast(em2, "revpairwise", infer = c(TRUE, TRUE))
data.frame(c2) |>
  kable(digits = 3, caption = "TEE RADPAD Contrast on Box-Cox Scale") |>
  kable_styling()
```

\newpage

## Anesthesia

### Observed Difference in Median Dose

```{r}
rA |>
  group_by(Procedure_Type) |>
  summarise(NODPAD = median(ifelse(RADPAD == "N", Dose, NA), na.rm = TRUE),
            RADPAD = median(ifelse(RADPAD == "Y", Dose, NA), na.rm = TRUE)) |>
  mutate(Difference = RADPAD - NODPAD) |> 
  mutate(Perecent = Difference/NODPAD*100) |>
  kable(digits = 2, caption = "Median Dose (microSV) Statistics for Anesthesia",
        col.names = c("Procedure", "Median NOPAD", "Median RADPAD",
                      "Difference", "Percent Change"))|>
  kable_styling()
```

### Gamma GLM Analysis of Variance

```{r}
## fit gamma GLM
fit_rA <- glm(Dose ~ RADPAD * Procedure_Type + Weight + Time,
             family = Gamma(link = "log"),
             data = rA)

Anova(fit_rA, type = "III") |>
  kable(digits = 3, caption = "Anesthesia Analysis of Variance")
```

### Contrasts 

```{r}
## RADPAD by Procedure Type effect on log scale
em0 <- emmeans(fit_rA, ~RADPAD|Procedure_Type)
c0 <- contrast(em0, "revpairwise", infer = c(TRUE, TRUE))
data.frame(c0) |>
  kable(digits = 3,
        caption = "Anesthesia RADPAD by Procedure Type Contrast on Log Scale") |>
  kable_styling()

## RADPAD by Procedure Type ratio
em <- emmeans(fit_rA, type = "response", ~RADPAD|Procedure_Type)
c1 <- contrast(em, method = "revpairwise", infer = c(TRUE, TRUE))
data.frame(c1) |>
  select(-df) |>
  kable(digits = 3,
        caption = "Anesthesia RADPAD by Procedure Type Contrast, Ratio of Geometric Means") |>
  kable_styling()
```

\newpage

```{r}
## RADPAD main effect on log scale
em2 <- emmeans(fit_rA, ~RADPAD)
c2 <- contrast(em2, "revpairwise", infer = c(TRUE, TRUE))
data.frame(c2) |>
  kable(digits = 3, caption = "Anesthesia RADPAD Contrast on Log Scale") |>
  kable_styling()

## RADPAD main effect ratio
em3 <- emmeans(fit_rA, type = "response", ~RADPAD)
c3 <- contrast(em3, method = "revpairwise", infer = c(TRUE, TRUE))
data.frame(c3) |>
  select(-df) |>
  kable(digits = 3,
        caption = "Anesthesia RADPAD Contrast, Ratio of Geometric Means") |>
  kable_styling()
```

\newpage 

## Resiudal Analysis

### Resident 1
```{r}
plot(fit_r1, which = c(1,2))
```

### Resident 2
```{r}
plot(fit_r2, which = c(1,2))
```

\newpage

### Faculty

```{r}
plot(fit_rF)
qqnorm(residuals(fit_rF, type = "pearson"))
qqline(residuals(fit_rF, type = "pearson"))
```

### Tee

```{r}
plot(fit_rTee)
qqnorm(residuals(fit_rTee, type = "pearson"))
qqline(residuals(fit_rTee, type = "pearson"))
```

\newpage

## Anesthesia

```{r, out.width = "100%", out.height="100%"}
plot(simulateResiduals(fit_rA))
```
```{r,  out.width = "75%", out.height="75%"}
plot(residuals(fit_rA, type = "deviance") ~ predict(fit_rA, type = "link"),
     xlab = expression(hat(eta)), ylab = "deviance residuals",
     main = "Fitted Values vs Deviance Residuals")
```


\newpage

## Numeric Summaries

### Dose (microSV)

```{r}
r |>
  group_by(Lab_Personnel, Procedure_Type, RADPAD) |>
  summarise(Count = n(),
            Mean = mean(Dose),
            Median = median(Dose),
            SD = sd(Dose),
            Min = min(Dose),
            Max = max(Dose)) |>
  kable(digits = 3) |>
  kable_styling()
```

$\space$

### Fluoroscopic Time (minutes)

```{r}
r |>
  group_by(Procedure_Type, RADPAD) |>
  summarise(Mean = mean(Time),
            Median = median(Time),
            SD = sd(Time),
            Min = min(Time),
            Max = max(Time)) |>
  kable(digits = 3) |>
  kable_styling()
```

$\space$

### Weight (kilograms)

```{r}
r |>
  group_by(Procedure_Type, RADPAD) |>
  summarise(Mean = mean(Weight),
            Median = median(Weight),
            SD = sd(Weight),
            Min = min(Weight),
            Max = max(Weight)) |>
  kable(digits = 3) |>
  kable_styling()
```

## Graphical Summaries

### Fluoroscopic Time (minutes)

```{r, out.width="100%", out.height="100%"}
ggplot(r, aes(RADPAD, Time, fill = RADPAD)) +
  geom_boxplot() +
  facet_wrap(vars(Procedure_Type)) +
  theme(legend.position="none")
```

### Weight (kilograms)

```{r, out.width="100%", out.height="100%"}
ggplot(r, aes(RADPAD, Weight, fill = RADPAD)) +
  geom_boxplot() +
  facet_wrap(vars(Procedure_Type)) +
  theme(legend.position="none")
```

