---
title: "RADPAD Analysis: Resident 1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(emmeans)
library(broom)
library(lme4)
options("scipen" = 3)
```

```{r}
r <- read_csv("radpad.csv") |>
  mutate(Relative_Dose = Dose/Patient_DAP)
r_R <- r |> filter(Occupation == "Resident 1")
```

## ANOVA for Relative Dose

**Model**: $\space RelativeDose \sim RADPAD + Procedure + (RADPAD \times Procedure) + Weight + Time$ 

where: $\space \space RelativeDose = \frac{Dose\space(microSV)}{PatientDAP}$

```{r}
m <- lm(Relative_Dose ~ RADPAD*Procedure + Weight + Time, data = r_R)
b <- MASS::boxcox(m, plotit=FALSE)
l <- b$x[which.max(b$y)]  ## log transform
m_R <- lm(log(Relative_Dose) ~ RADPAD*Procedure + Weight + Time, data = r_R)
anova(m_R)
```

## Contrast Estimates for RADPAD Effect on Log Scale

```{r}
rg <- ref_grid(m_R)
em <- emmeans(rg, ~RADPAD)
c <- contrast(em, interaction = "revpairwise")
knitr::kable(summary(em, infer = c(FALSE, FALSE)))
knitr::kable(tidy(summary(c, infer = c(TRUE, TRUE))))
```

## Contrast Estimates for RADPAD Effect on Response Scale

```{r}
rg <- ref_grid(m_R)
em <- emmeans(regrid(rg, transform = "response"), ~ RADPAD)
c <- contrast(em, interaction = "revpairwise")
knitr::kable(tidy(summary(em, infer = c(FALSE, FALSE))))
knitr::kable(tidy(summary(c, infer = FALSE)))
p_red <- abs(round((tidy(em)$response[2]-tidy(em)$response[1])/tidy(em)$response[1]*100, 2))
```

`r p_red`% reduction in Relative Dose for Resident 1 due to RADPAD, averaging over Procedure Type and controlling for Weight and Time.

\newpage

## Contrast Estimates for RADPAD by Procedure Interaction on Log Scale

```{r}
rg <- ref_grid(m_R)
em <- emmeans(rg, ~RADPAD|Procedure)
c <- contrast(em, interaction = "revpairwise")
sumem <- tidy(summary(em, infer = FALSE))
knitr::kable(sumem)
sumc <- tidy(summary(c, infer = TRUE)) |>
  rename("RADPAD  " = RADPAD_revpairwise)
knitr::kable(sumc)
```

## Contrast Estimates for RADPAD by Procedure Interaction on Response Scale

```{r}
rg <- ref_grid(m_R)
em <- emmeans(regrid(rg, transform = "response"), ~ RADPAD|Procedure)
c <- contrast(em, interaction = "revpairwise")
knitr::kable(tidy(summary(em, infer = c(FALSE, FALSE)), digits = 10))
knitr::kable(tidy(summary(c, infer = FALSE)), digits = 10)
```



```{r}
##shapiro.test(residuals(m_R))
```


```{r}
##car::ncvTest(m_R)
```

```{r}
##lmtest::dwtest(m_R)
```

## Residual Analysis

```{r, out.width = "50%", out.height = "50%"}
plot(m_R)
```

## Influential Data Points

```{r}
r_o <- r_R[c(85, 156, 168, 172, 189, 191), ] |>
  select(MRN, Occupation, Procedure, Weight, Time, Dose, Patient_DAP, Relative_Dose)
knitr::kable(r_o)
```





